An Early Eocene temperature gradient modelled with geological and ecological proxies

### 2022-10-13 manuscript/manuscript.Rmd
in-line code for global average T in manuscript

@@

The global Eocene mean sea surface temperature is estimated at 28.8 (95% CI: 25.7 - 31.7) degC, significantly higher than the modern (?? degC).

>>>>

```{r results block 1 - for in-line code}
setwd('..') # move up to parent folder of /manuscript
source("R/functions/model_components/gradient.R")
mode_all <- readRDS("results/eeco/eeco_climate_model_output_params.rds")
modm_all <- readRDS("results/modern/modern_climate_model_output.rds")

# latitudinal weights in 1 deg lat bands
alpha1 <- seq(1,90,1)
alpha2 <- seq(0,89,1)
latweight <- sin(pi*alpha1/180) - sin(pi*alpha2/180) # sums to 1 - careful when using lat. subsets
# eocene temperature gradient
eotemp <- gradient(seq(0.5,89.5,1), mode_all[,1:4],0)
# eocene temperature gradient quantiles
eotempq <- apply(eotemp,2,function(x) quantile(x, probs = c(0.025,0.5,0.975)))
# eocene global average temperature with 95% CI
eotemp_global <- apply(eotempq,1, function(x) round(sum(x*latweight),1), simplify = FALSE)
# modern temperature gradient
motemp <- gradient(seq(0.5,89.5,1), modm_all[,1:4],0)
# modern temperature gradient quantiles
motempq <- apply(motemp,2,function(x) quantile(x, probs = c(0.025,0.5,0.975)))
# modern global average temperature with 95% CI
motemp_global <- apply(motempq,1, function(x) round(sum(x*latweight),1), simplify = FALSE)

```

@@

The global Eocene mean sea surface temperature is estimated at `r eotemp_global$"50%"` (95% CI: `r eotemp_global$"2.5%"` - `r eotemp_global$"97.5%"`) degC, significantly higher than the modern (`r motemp_global$"50%"` degC).
